---
title: "Package dataframeinfo vignette"
author: "Bolim Son"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{my-vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(dataframeinfo)
```

### __Description__

The package `dataframeinfo()` is an package to explore a dataframe before going to an analysis.

There are many ways to inspect a dataframe, but it is quite cumbersome to write each function.

Hence, in this package, function `info()` just does the multiple functions at once and generates a list of information.

This package was made because the author had spent hours searching for the wrong fixes.


### __Function `info()`__

Function `info()` has two arguments

* `x` is a data frame. Input format must be a data frame. If not, the function will not run though. Use `data.frame()` or other functions to coerce your `x` into a data frame.

* `y` is a column name in the specified data frame `x`. `y` must be a column name in `x`. If not, the function will not run through. Use `colnames(data_frame_name)` to first inspect your columns and write the column of your interest at `y`. Use `""` to write the column name. 

* Example
```{r}
# Example 1

# Arbitrary data frame generated for x
# column name specified

info(x = data.frame(a = rnorm(10, mean = 5, sd=1), b = 1:10), y = "a")
```

```{r}
# Example 2

# Using data frame in R
info(x = mtcars, y = "cyl")

```

```{r}
# Example 3

# Using data frame in R

info(x = iris, y = "Petal.Length")
```

### __Errors__

The function `info()` will not run through if it encounters errors. 

* `x` needs to be a data frame

```{r, error = TRUE}
# Error example 1

info(x = list(a = rnorm(10, mean = 5, sd=1), b = 1:10), y = "a")

```

* `y` needs to be a column name of  `x`

```{r, error = TRUE}
# Error example 2

info(x = data.frame(a = rnorm(10, mean = 5, sd=1), b = 1:10), y = "c")

```

