---
title: "GR5072 Modern Data Structures_HW 9"
author: "Bolim Son"
date: "November 20, 2019"
output:
  html_document:
    df_print: paged
    toc: no
  html_notebook:
    df_print: paged
    toc: no
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
# Load libraries
library(httr)
library(rvest)
library(dplyr)
library(stringr)

```


#### __1. Get HTML__

Get the content of the page into R.

```{r}
# Set url for Mitch McConnell's Wikipedia page
url <- "https://en.wikipedia.org/wiki/Mitch_McConnell"

# Get content
wiki <- read_html(url)

# Check the structure of content retrieved
str(wiki, max.level = 1)

```



#### __2. Get the info box__

On the right side of the page is a box of structured content, called an info box. Wikipedia has many types of such info boxes to provide content comparably for a group of articles of the same class (e.g. the Members of the U.S. senate, Fortune 500 companies, Crime Syndicates etc.)

a) Find the CSS class of the infobox.
```{r}
infobox_node <- html_nodes(wiki, "table.infobox")

```


b) Extract the part of the HTML document that contains the infobox using the CSS information.
```{r}
infobox <- html_table(infobox_node)

# Inspect the extracted infobox
str(infobox)
```



#### __3. Make a data frame__

a) Parse the infobox table HTML you obtained above into a data frame.

```{r}
# Parse to a data frame
infobox_df <- as.data.frame(infobox[[1]])

# Check the data frame
head(infobox_df)
```


b) Name the columns of the table you obtain key and value. So, in the example for Mitch McConnell, "Deputy" would be the key, and the content information (i.e. the value) is "John Cornyn".

```{r}
# Place name as the first row
infobox_df[1,] <- colnames(infobox_df)

# Set column names
colnames(infobox_df) <- c("key", "value")

# Check infobox_df
head(infobox_df)
```


c) Filter the data frame (and rename variables if necessary) to the "Full name", "Political Party", and "Children". Use this selection of variables for all subsequent questions.

```{r}
# Rename variable
infobox_df[1,1] <- "Full name"

# Filter using dplyr
# Set variables
vars <- c("Full name", "Political party", "Children")
# Filter out the variables
infobox_filter <- infobox_df %>%
  filter(key %in% vars)

# Show the filtered result
infobox_filter
```



#### __4. Make a function__

a) Use the code above to make a function called get_wiki_info that uses a single input url (a Wikipedia URL) and outputs the data frame of the format above. There is no need to account for exceptions (e.g. no info box on the page; page does not exist etc.) - we will only use members of the U.S. Senate for this exercise.

```{r}
get_wiki_info <- function(url = "https://en.wikipedia.org/wiki/Mitch_McConnell") { # Set default value as Mitch McConnell's wikipedia page

# Retrieve a data.frame from wikipedia  
  df <- read_html(url) %>%
    html_nodes(., "table.infobox") %>%
    html_table(.)
  
  df <- as.data.frame(df[[1]])
# Rename some variables in the data.frame  
  df[1, ] <- colnames(df)
  df[1,1] <- "Full name"
  colnames(df) <- c("key", "value")
# Set variables of selection  
  variables <- c("Full name", "Political party", "Children")
# Filter the variables  
  filter_df <- df %>%
    filter(key %in% variables)
  
# Return result  
  return(filter_df)
}


```


b) Show how your function works on the following two URLs:

https://en.wikipedia.org/wiki/Tammy_Duckworth , https://en.wikipedia.org/wiki/Susan_Collins
Depending on your previous function, you may receive an error message because Susan Collins has no entry for Children. Fix your function so that NA is recorded in such instances.

```{r}
# Run get_wiki_info with Tammy Duckworth
get_wiki_info(url = "https://en.wikipedia.org/wiki/Tammy_Duckworth")

```

```{r}
# Run get_wiki_info with Susan Collins
get_wiki_info(url = "https://en.wikipedia.org/wiki/Susan_Collins")
```

Results for Susan Collins does not have 'Children' variable. To record NA for children, the function will be modified.

```{r}
get_wiki_info <- function(url = "https://en.wikipedia.org/wiki/Mitch_McConnell") { # Set default value as Mitch McConnell's wikipedia page

# Retrieve a data.frame from wikipedia  
  df <- read_html(url) %>%
    html_nodes(., "table.infobox") %>%
    html_table(.)
  
  df <- as.data.frame(df[[1]])
# Rename some variables in the data.frame  
  df[1, ] <- colnames(df)
  df[1,1] <- "Full name"
  colnames(df) <- c("key", "value")
# Set variables of selection  
  variables <- c("Full name", "Political party", "Children")
# Filter the variables  
  filter_df <- df %>%
    filter(key %in% variables)

# Check for Children
  if("Children" %in% filter_df$key == FALSE){
    edit <- tibble::add_row(filter_df, key = 'Children', value = NA) # Add a row
    return(edit)
    
  } else {
    return(filter_df) # If have children, just return the dataframe
  }
  
}
```


```{r}
# Run edited function with Susan Collins
get_wiki_info(url = "https://en.wikipedia.org/wiki/Susan_Collins")
```



#### __5. Get all senators' pages (Optional)__

a) On this page (https://en.wikipedia.org/wiki/Current_members_of_the_United_States_Senate) you find a list of all current senators of the U.S. congress.
Import the site and obtain a vector with the URLs for the Wikipedia sites of all 100 members of congress (hint: the function xml_attr is one option). Note, this should only include the URLs and remove all other content.

```{r, cache = TRUE}
## Set r chunck set as cache = TRUE and save retrieved data when knitting


# Write a for loop to extract senator vectors from Wikipedia page

# Set i as 2 because that is the first starting row
i <- 2

# Create a data.frame to store senator vector
senator_names_df <- data.frame(names = 1:100)

# Set a senator_url for extraction
senator_url <- "https://en.wikipedia.org/wiki/Current_members_of_the_United_States_Senate"


# For loop
for (i in i:101) {

# Make the loop run until the 100th one
  if (i < 102) {
    xpath_loop <- str_c('//*[@id="senators"]/tbody/tr[', i, ']/th')
    
    senator_names_df$names[i-1] <- read_html(senator_url) %>% 
    html_node(xpath = xpath_loop) %>% 
    html_node("a") %>% 
    xml_attr('href') %>% # Choose attribute
    str_c(.)  # Change to a string

    senator_names_df
        
    } else {
      return(senator_names_df) # Return data.frame
    }
}


# Check data frame
head(senator_names_df)

```


b) Create a loop that uses your get_wiki_info function to import the information on political party and number of children for all members of the senate.

```{r, cache= TRUE}
## Set r chunck set as cache = TRUE and save retrieved data when knitting

# Create a list to place senator information
senator_info <- list()

i <- 1
for (i in i:100){
  
  if(i < 101){
# Set url for retrieving information  
  senator_url <- str_c("https://en.wikipedia.org", senator_names_df$names[i])

# Save the information in a list, extract only the values wanted
  senator_info[[i]] <- t(get_wiki_info(url = senator_url))[2,]
}}

# Bind it all together
senator_info_df = do.call(rbind, senator_info)
senator_info_df <- data.frame(senator_info_df)
colnames(senator_info_df) <- c("Full Name", "Political_party", "Children")

# Check the binded dataframe
head(senator_info_df)

```


c) Clean up your data to provide some summary statistics. How many children does the average senator have? What about by party affiliation?

```{r}
# Clean the data
senator_info_clean <- senator_info_df %>%
# Clean the political party
  mutate(Political_party = 
           str_remove(senator_info_df$Political_party, " *\\(.*\\)$")) %>%
# Clean children
  mutate(Children = 
           str_remove_all(senator_info_df$Children, 
                          "\\(, *including .*|son|daughter|1stepchild")) %>%
  mutate(Children = 
           str_remove_all(senator_info_df$Children, "\\[3\\]\\[4\\]")) 

# Clean children
senator_info_clean <- senator_info_clean %>%
  mutate(Children = 
           str_remove_all(senator_info_clean$Children, "\\s")) %>%
  mutate(Children =
           str_replace_all(senator_info_clean$Children, "[A-z]{1,}", "1")) %>%
  mutate(Children = as.numeric(Children)) 

# Check cleaned data
head(senator_info_clean)
```


```{r}
# Average children a senator has, among the ones that have children
mean(senator_info_clean$Children, na.rm = TRUE)

# Average children a senator has, if NA is counted as 0
children_all <- if_else(is.na(senator_info_clean$Children), 0, senator_info_clean$Children)
mean(children_all)

```

Average children a senator has, among the onese that have children, are approximately 2.77.

If senators without children are included, on average, a senator has 2.05 children.


```{r}
# Party affiliation
senator_info_clean %>%
  group_by(Political_party) %>%
  summarise(count = n())

```

There are 53 Republican senators, 46 Democratic senators, and 1 indpendent senator.