---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# dataframeinfo

<!-- badges: start -->
<!-- badges: end -->

The goal of dataframeinfo is to ...
inspect a variable inside a data set before going into analysis.


## Installation

You can install the released version of dataframeinfo from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("dataframeinfo")
```

And the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("QMSS-G5072-2019/Son_Bolim")
```
## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(dataframeinfo)
## basic example code

## This is how you use it
## For x, enter a data frame
## For y, enter a variable name inside "" like "variable_name"
## The function will generate a list of some basic information about that variable.

info(x = data.frame(a = rnorm(1:10), b = rnorm(11:20)), y = "b")

```


This will return errors:

```{r, error = TRUE, eval = FALSE}
# x needs to be in a data frame format.
# y has to be a column name in the data frame x specified.

info(x = list(a = c(1:5), b = c(11:15)), y = "b")
```

Change x into a data frame:

```{r}
# x is changed to a data frame

info(x = data.frame(a = 1:5, b = 11:15), y = "b")

```

